Comparison of Catalogs

### 2024-09-22 compare/.ipynb_checkpoints/compare-checkpoint.qmd
- catalog completeness (dis and T) - pairplot with correlation - styling

---
title: "Comparison of Catalogs"
format: html
toc: true
toc-depth: 2
theme: minti
bibliography: ../My_Library.bib
execute: 
  echo: false
  eval: true
  warning: false
---

@@

-   The data have been joined based on their position in the sky (Ra, Dec, Distance).
-   We use TOPCAT to create two joins, an inner and an outer join
-   We will use the inner join for 1-1 comparisons
-   If we see that the data are similar we can use the outer join
-   For the comparison we keep the parameters names exactly they are given in the catalogs

```{python}
#| echo: false
import numpy as np
import matplotlib.pyplot as plt
from sklearn.metrics import r2_score
from astropy.io import ascii
from astropy.coordinates import SkyCoord
from astropy.table import Table, join, QTable
import astropy.units as u
from astropy.visualization import quantity_support, hist
#quantity_support()
from astropy.stats import sigma_clip, SigmaClip
from astropy.modeling import models, fitting
from scipy.stats import pearsonr
import seaborn as sns
import pandas as pd
import glob
import os
plt.style.use('ggplot')
pd.set_option('display.float_format', lambda x: '%.f' % x)

dt = QTable(ascii.read("../tables/inner_join.ecsv"), masked=True)
```

The dataset we are going to use consists of `{python} len(dt)` galaxies and `{python} len(dt.colnames)` columns.

@@

    -   Correlation Heatmaps: A correlation heatmap is a graphical tool that displays the correlation between multiple variables as a color-coded matrix. It’s like a color chart that shows us how closely related different variables are. In a correlation heatmap, each variable is represented by a row and a column, and the cells show the correlation between them. The color of each cell represents the strength and direction of the correlation, with darker colors indicating stronger correlations.

@@

::: panel-tabset
## Right Ascension

@@

## Declination

@@

## Distance

@@

::: panel-tabset
## Heliocentric radial Velocity

@@

## Scatter Grid

```{python}
vel_data = dt[["RVel", "V", "VLG", "cz", "V_VIR"]].to_pandas()

# Create a pair grid
sns.pairplot(vel_data, corner=True)
plt.show()
```

## Heatmap

```{python}
# Compute the correlation matrix
corr_matrix = vel_data.corr()

# Create a heatmap
plt.figure(figsize=(10, 8))
sns.heatmap(corr_matrix, annot=True, cmap="coolwarm", vmin=-1, vmax=1)
plt.title("Correlation Heatmap of Velocity Values")
plt.show()
```
:::

@@

::: panel-tabset
## Galaxy Types

```{=html}
<table style="border-collapse:collapse;border-spacing:0" class="tg"><thead>
<tr><th style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;font-weight:normal;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">Hubble stage T </th><th style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;font-weight:normal;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">−6 </th><th style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;font-weight:normal;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">−5 </th><th style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;font-weight:normal;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">−4 </th>
<th style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;font-weight:normal;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">−3 </th><th style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;font-weight:normal;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">−2 </th><th style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;font-weight:normal;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">−1 </th><th style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;font-weight:normal;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">0 </th>
<th style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;font-weight:normal;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">1 </th><th style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;font-weight:normal;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">2 </th><th style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;font-weight:normal;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">3 </th><th style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;font-weight:normal;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">4 </th>
<th style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;font-weight:normal;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">5 </th><th style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;font-weight:normal;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">6 </th><th style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;font-weight:normal;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">7 </th><th style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;font-weight:normal;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">8 </th>
<th style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;font-weight:normal;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">9 </th><th style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;font-weight:normal;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">10 </th><th style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;font-weight:normal;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">11 </th></tr>
</thead>
<tbody>
<tr><td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">de Vaucouleurs class </td><td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">cE </td><td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">E </td><td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">E+ </td>
<td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">S0− </td><td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">S00 </td><td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">S0+ </td><td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">S0/a </td>
<td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">Sa </td><td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">Sab </td><td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">Sb </td><td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">Sbc </td>
<td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">Sc </td><td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">Scd </td><td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">Sd </td><td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">Sdm </td>
<td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">Sm </td><td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">Im </td><td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal"> </td></tr>
<tr><td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">Approximate Hubble class </td><td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal" colspan="3">E </td><td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal" colspan="3">S0 </td><td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">S0/a </td>
<td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">Sa </td><td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">Sa-b </td><td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">Sb </td><td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">Sb-c </td>
<td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal" colspan="3">Sc </td><td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">Sc-Irr </td><td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal">Irr I </td><td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal"></td>
<td style="border-color:inherit;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;text-align:center;vertical-align:top;word-break:normal"></td></tr>
</tbody></table>

@@

scatter_plot(dt["TType"], dt["T"], yerr = dt["E_T"])

@@

dt["diff_T"] = relative_diff(hec = dt["T"], lcv = dt["TType"])
percent_desc_histogram(dt["diff_T"], zoom = [-200,200])
```

[?] After the sigma clip we only lose 39 galaxies ($14\%$) and we can see that both the median and the mean of the percentage change are close to $0\%$.This is why we can assume that the Types of the galaxies are the same for the two catalogs

## Inclination

```{python}
temp = dt[["inc", "INCL"]].to_pandas()

@@

## Major Axis

@@

```{python}
#| echo: false
logKLum_corr = round(compare_data(dt["logKLum"], dt["logL_K"], sigma = True)[3], 3)

@@

```{python}
#| echo: false
mag_B_corr = round(compare_data(dt["mag_B"], dt["BT"], sigma = True)[3], 3)
Kmag_corr = round(compare_data(dt["Kmag"], dt["K"], sigma = True)[3], 3)
```

@@

::: panel-tabset
## B mag

@@

## K mag

scatter_plot(dt["Kmag"], dt["K"], yerr = dt["E_K"], sigma = True)
```

@@

[?]
:::

## Absorption

```{python}
#| echo: false
ab_corr = round(compare_data(dt["AB"], dt["AG"])[3], 3)
#ab_int_corr = round(compare_data(dt["AB_int"], dt["AI"])[3], 3)
```

|  LCV   | HECATE |                   Description                   | Pearson Correlation \[-1,1\] |
|:----------------:|:----------------:|:-----------------:|:----------------:|
|   AB   |   AG   |    Galactic extinction/absorption in B band     |      `{python} ab_corr`      |
| AB_int |   AI   | Internal/Intrinsic B band extinction/absorption |              0               |

::: panel-tabset
## Galactic

```{python}
scatter_plot(dt["AB"], dt["AG"], logscale = True)#, scatter_plot(dt["AB_int"], dt["AI"])
```

## Internal
```{python}
sns.jointplot(x=dt["AB_int"], y=dt["AI"], kind="reg",)
plt.xlabel(f"LCV, AB_int [{dt['AB_int'].unit}]")
plt.ylabel(f"HECATE, AI [{dt['AI'].unit}]")
plt.show()
```
Here we can see an attempt to correlate the internal absorption of the two catalogs, however as it is very clear, there is no correlation, since even their distributions are not similar.
[?]
:::

@@

::: panel-tabset
## Heatmap

```{python}
#| echo: false
sfr_data.drop("logSFR_GSW", axis=1, inplace=True)
# Compute the correlation matrix
corr_matrix = sfr_data.corr()

# Create a heatmap
plt.figure(figsize=(10, 8))
sns.heatmap(corr_matrix, annot=True, cmap="coolwarm", vmin=-1, vmax=1)
plt.title("Correlation Heatmap of SFR Values")
plt.show()
```

## Scatter Grid

```{python}
# Create a pair grid 
plt.figure(figsize=(10, 8))
sns.pairplot(sfr_data, corner=True)
plt.show()
```

## Density Grids

```{python}
# Create a pair grid 
sns.pairplot(sfr_data, corner=True, kind="kde", diag_kind="hist")
plt.show()
```

:::

[?] It is possible we dont see a good correlation because we dont have a big enough common sample

@@

::: panel-tabset
## Stellar Masses Comparison

@@

dt["diff_M"] =  relative_diff(hec = dt["logM_HEC"].data, lcv = dt["logStellarMass"])
temp = dt[["diff_M"]].to_pandas()
sns.histplot(temp, kde = True)
plt.title("Percentage change Distribution [%]")
plt.xlabel(r"$\log_{10}M_*$ "+f"[{dt['logM_HEC'].unit}]")
plt.show()

@@

```{python}
temp.rename(columns={"diff_M":"Percantage Change [%]"}, inplace = True)
```
```{python}
temp.describe(percentiles=[])
```
## Heatmap

```{python}
mass_data.drop("logM_GSW", axis=1, inplace=True)
# Compute the correlation matrix
mass_corr_matrix = mass_data.corr()

# Create a heatmap
sns.heatmap(mass_corr_matrix, annot=True, cmap="coolwarm", vmin=-1, vmax=1)
plt.title("Correlation Heatmap of Mass Values")
plt.show()
```

## Plot Grid

```{python}
# Create a pair grid
sns.pairplot(mass_data, corner=True)
plt.show()
```

## Density Grid

```{python}
# Create a pair grid
sns.pairplot(mass_data, corner=True, kind="kde", diag_kind="hist")
plt.show()
```
:::

>>>>

---
title: "Comparison of Catalogs"
format: 
  html: 
    code-fold: true
    html-math-method: katex
  pdf: default
  arxiv-pdf:
    keep-tex: true    
toc: true
toc-depth: 3
theme: minti
bibliography: ../My_Library.bib
execute: 
  echo: false
  eval: true
  warning: false
---

@@

-   The data have been joined based on their position in the sky (Ra, Dec).
    -   We assume that every galaxy within 2 arc seconds of the initial coordinates is the same galaxy.
-   We use TOPCAT to create two joins, an inner and an outer join
-   We will use the inner join for 1-1 comparisons
-   If we see that the data are similar we can use the outer join
-   For the comparison we keep the parameters names exactly they are given in the catalogs

```{python}
#| echo: false
import numpy as np
import matplotlib.pyplot as plt
from sklearn.metrics import r2_score
import astropy
from astropy.io import ascii
from astropy.coordinates import SkyCoord
from astropy.table import Table, join, QTable
import astropy.units as u
from astropy.visualization import quantity_support, hist
#quantity_support()
from astropy.stats import sigma_clip, SigmaClip
from astropy.modeling import models, fitting
from scipy.stats import pearsonr
import seaborn as sns
import pandas as pd
import glob
import os
from tabulate import tabulate

plt.style.use('ggplot')
pd.set_option('display.float_format', lambda x: '%.f' % x)

dt = QTable(ascii.read("../tables/inner_join.ecsv"), masked=True)

inner = dt
outer = QTable(ascii.read("../tables/outer_join.ecsv"), masked=True)
hec_not_lcv = QTable(ascii.read("../tables/HEC_not_LCV_join.ecsv"), masked=True)
lcv_not_hec = QTable(ascii.read("../tables/LCV_not_HEC_join.ecsv"), masked=True)

lcv = QTable(ascii.read("../tables/final_table.ecsv"), masked=True)
hec = QTable(ascii.read("../tables/HECATE_LCV.ecsv"), masked=True)
```

The dataset we are going to use for the comparison (inner join) consists of `{python} len(dt)` galaxies and `{python} len(dt.colnames)` columns.

# Catalog Completeness

Checking for completeness in galaxy catalogs is essential to ensure that the data accurately represents the true population of galaxies. Incomplete catalogs can lead to biased results in statistical studies, such as the distribution of galaxy luminosity, mass, or star formation rates. Additionally, missing galaxies, especially those at faint magnitudes or large distances, can distort cosmological measurements and hinder our understanding of galaxy formation and evolution.

Completeness checks are crucial for addressing selection biases, ensuring accurate redshift distributions, and validating galaxy simulations. They help identify gaps in the data and guide follow-up observations, ensuring that the catalog provides a reliable sample for scientific analysis. Without these checks, conclusions drawn from the data may be inaccurate or incomplete.

------------------------------------------------------------------------

Checking for completeness in galaxy catalogs is essential to ensure that the data accurately represents the true population of galaxies. Incomplete catalogs can lead to biased results in statistical studies, such as the distribution of galaxy luminosity, mass, or star formation rates. Additionally, missing galaxies, especially those at faint magnitudes or large distances, can distort cosmological measurements and hinder our understanding of galaxy formation and evolution. Completeness checks are crucial for addressing selection biases, ensuring accurate redshift distributions, and validating galaxy simulations.

Distance-based corrections are applied to mitigate these biases by adjusting for the underrepresentation of galaxies at greater distances. As galaxies move farther away, they become fainter and harder to detect, leading to a drop in the number of detected galaxies. Methods like **volume corrections** (e.g., V/Vmax) and **luminosity function-based corrections** help account for these effects by estimating the true galaxy population based on the observed sample. These corrections ensure that statistical analyses, even in incomplete catalogs, more accurately reflect the full galaxy population.

```{python}
#| output: asis
# Creating a dataframe with the lengths of each table
data = {
    "Table": ["Inner join", "Outer join", "LCV", "HECATE", "Unique galaxies in LCV", "Unique Galaxies in Hecate"],
    "Number of galaxies": [len(dt), len(outer), len(lcv), len(hec), len(lcv_not_hec), len(hec_not_lcv)]
}

df_lengths = pd.DataFrame(data)

# Pretty print the dataframe in markdown format
df_lengths_md = df_lengths.to_markdown(index=False)
print(df_lengths_md)

## Completeness of the Inner join

$$
\text{Completeness (X)}=\frac{\text{(Galaxies in Inner Join)}}{\text{(Galaxies in X)}}×100\%
$$

Completeness (HECATE)= `{python} round(len(dt)/len(hec) *100)` %

Completeness (LCV)= `{python} round(len(dt)/len(lcv) *100)` %

## Completeness in Outer join

$$
\text{Completeness (X)}=\frac{\text{(Galaxies in Outer Join form X)}}{\text{(Galaxies in X)}}×100\%
$$

Completeness (HECATE)= `{python} round(len(hec_not_lcv)/len(hec)*100)` %

Completeness (LCV)= `{python} round(len(lcv_not_hec)/len(lcv)*100)` %

Combined Completeness =$\frac{\text{Total galaxies in Outer}}{\text{Unique galaxies in HECATE + LCV}}$= `{python} round(len(outer)/(len(lcv_not_hec)+(len(hec_not_lcv))) * 100)` %

## Completeness of the Catalogs, based on the Distance and the Morphological Type

```{python}
#| fig-cap: "Histograms showing the Completeness of the Catalogs"
#| layout-ncol: 2
#| fig-subcap: 
#| - "HECATE"
#| - "LCV"
#| label: fig-dis-comp

plt.close()
# Using numpy's histogram function to get the bin edges automatically for the first dataset
_, bin_edges = np.histogram(hec["D"].data, bins='auto')  # 'auto' decides the best number of bins for the data

# Step 2: Use the same bin edges for all datasets
sns.histplot(hec["D"].value, bins=bin_edges, kde=True, label="HECATE")
sns.histplot(hec_not_lcv["D"].value, bins=bin_edges, kde=True, label="Unique for Hecate")
sns.histplot(inner["D"].value, bins=bin_edges, kde=True, label="Inner join")

plt.title("Completeness of Distance for HECATE")
plt.xlabel("Distance [Mpc]")
plt.ylabel("Number of galaxies")
plt.legend()
plt.show()
plt.close()

# Using numpy's histogram function to get the bin edges automatically for the first dataset
_, bin_edges = np.histogram(lcv["Dis"].data, bins='auto')  # 'auto' decides the best number of bins for the data
sns.histplot(lcv["Dis"].value, bins=bin_edges, kde = True, label = "LCV")
sns.histplot(lcv_not_hec["Dis"].value, bins=bin_edges, kde = True, label = "Unique for LCV")
sns.histplot(inner["Dis"].value, bins=bin_edges, kde = True, label = "Inner join")
plt.title("Completeness of Distance for LCV")
plt.xlabel("Distance [Mpc]")
plt.ylabel("Number of galaxies")
plt.legend()
plt.show()
plt.close()

```{python}
#| layout-ncol: 2
#| fig-cap: "Histograms showing the Completeness of the Catalogs"
#| fig-subcap: 
#| - "HECATE"
#| - "LCV"
#| label: fig-type-comp
# Using numpy's histogram function to get the bin edges automatically for the first dataset
bin_edges = 16

sns.histplot(hec["T"], bins=bin_edges, kde = True, label = "HECATE")
sns.histplot(hec_not_lcv["T"], bins=bin_edges, kde = True, label = "Unique for Hecate")
sns.histplot(inner["T"].value, bins=bin_edges, kde = True, label = "Inner join")
plt.title("Completeness of Types of galaxies for HECATE")
plt.xlabel("Type")
plt.ylabel("Number of galaxies")
plt.legend()
plt.show()
plt.close()

bin_edges = 14

sns.histplot(lcv["TType"], bins=bin_edges, kde = True, label = "LCV")
sns.histplot(lcv_not_hec["TType"], bins=bin_edges, kde = True, label = "Unique for LCV")
sns.histplot(inner["TType"].value, bins=bin_edges, kde = True, label = "Inner join")
plt.title("Completeness of Types of galaxies for LCV")
plt.xlabel("Type")
plt.ylabel("Number of galaxies")
plt.legend()
plt.show()
plt.close()

As we can see from the histograms @fig-dis-comp and @fig-type-comp the sample of nique galaxies of each catalog, gets smaller by an almost constant proportion (Inner join).

This means there is no bias in the selection of the galaxies.

@@

<!-- -->

-   Correlation Heatmaps: A correlation heatmap is a graphical tool that displays the correlation between multiple variables as a color-coded matrix. It's like a color chart that shows us how closely related different variables are. In a correlation heatmap, each variable is represented by a row and a column, and the cells show the correlation between them. The color of each cell represents the strength and direction of the correlation, with darker colors indicating stronger correlations.

# Define a function that creates a pairplot with correlation coefficients
def pairplot_with_correlation(df, cmap="coolwarm", font_size=12, font_style='italic', font_weight='bold', box_color='white'):
    """
    Creates a Seaborn pairplot with correlation coefficients displayed on the upper triangle.

    Parameters:
    df : pandas.DataFrame
        The dataframe containing the data for the pairplot.
    cmap : str, optional
        The colormap to use for coloring the correlation coefficients (default is 'coolwarm_r').
    font_size : int, optional
        The font size for the correlation coefficient annotations (default is 12).
    font_style : str, optional
        The font style for the correlation coefficient annotations (default is 'italic').
    font_weight : str, optional
        The font weight for the correlation coefficient annotations (default is 'bold').
    box_color : str, optional
        The background color of the box around the correlation coefficients (default is 'white').

    Returns:
    A Seaborn pairplot with correlation coefficients displayed in a heatmap style.
    """
    plt.close()
    # Create Pairplot
    pairplot = sns.pairplot(df, plot_kws={"color": "green"}, diag_kws = {"color" : "green"})

    # Calculate Correlations
    corr_matrix = df.corr()

    # Define a colormap
    colormap = plt.get_cmap(cmap)

    # Overlay Correlation Coefficients with customized annotations
    for i, j in zip(*np.triu_indices_from(corr_matrix, 1)):
        # Get the correlation value
        corr_value = corr_matrix.iloc[i, j]

        # Get a color from the colormap based on the correlation value
        color = colormap((corr_value + 1) / 2)  # Normalize between 0 and 1

        # Annotate the plot with the correlation value, using a box, bold, and italic
        pairplot.axes[i, j].annotate(f"{corr_value:.2f}", 
                                     xy=(0.5, 0.5), 
                                     xycoords='axes fraction',
                                     ha='center', 
                                     va='center', 
                                     fontsize=font_size,  # Font size
                                     fontstyle=font_style,  # Font style
                                     fontweight=font_weight,  # Font weight
                                     bbox=dict(facecolor=box_color, edgecolor=color, boxstyle='round,pad=0.5'),  # Box around the text
                                     color=color)  # Use heatmap color for text

    # Show the plot
    plt.show()

@@

::: panel-tabset
### Right Ascension

@@

### Declination

@@

### Distance

@@

@@

```{python}
vel_data = dt[["RVel", "V", "VLG", "cz", "V_VIR"]].to_pandas()
```

```{python}
pairplot_with_correlation(vel_data)
```

@@

::: panel-tabset
### Galaxy Types

@@

scatter_plot(dt["TType"], dt["T"], yerr = dt["E_T"])
```

@@

dt["diff_T"] = relative_diff(hec = dt["T"], lcv = dt["TType"])
percent_desc_histogram(dt["diff_T"], zoom = [-200,200])
dt["diff_T_clip"] = sigma_clip(dt["diff_T"], sigma=3)
dt[["diff_T", "diff_T_clip"]].to_pandas().describe()

\[?\] After the sigma clip we only lose 39 galaxies ($14\%$) and we can see that both the median and the mean of the percentage change are close to $0\%$.This is why we can assume that the Types of the galaxies are the same for the two catalogs

#### Normalize the scale of galaxy types

It is very possible that the two catalogs use different scaling methods, as indicated by the use of decimal numbers in HECATE.

```{python}
types = dt[["T", "TType"]].to_pandas()
```

```{python}
def min_max_normalize(series):
    return (series - series.min()) / (series.max() - series.min())

types["T_norm"] = min_max_normalize(types["T"])
types["T_norm"].mask = types["T"].mask
types["TType_norm"] = min_max_normalize(types["TType"])
types.describe()
```

Also, as we can see the minimum values are lower by 2 in HECATE, which complies with the linear fit.

```{python}
sns.histplot(types[["TType_norm", "T_norm"]], kde = True)
plt.xlabel("Types of galaxies")
plt.show()
```

```{python}
types["norm_diff"] = (-types["TType_norm"] + types["T_norm"])
sns.histplot(types["norm_diff"], kde = True)

plt.xlabel(r"$T_{HECATE}-T_{LCV}$")
plt.show()
```

```{python}
types["norm_rel_diff"] = (-types["TType_norm"] + types["T_norm"])/types["T"]
sns.histplot(types["norm_rel_diff"], kde = True)
plt.xlim(-0.2,0.2)
plt.xlabel(r"$(T_{HECATE}-T_{LCV})/T_{HECATE}$")
plt.show()
```

### Inclination

temp = dt[["inc", "INCL"]].to_pandas()

@@

### Major Axis

@@

```{python}
#| echo: false
#| warning: false
logKLum_corr = round(compare_data(dt["logKLum"], dt["logL_K"], sigma = True)[3], 3)

@@

```{python}
#| echo: false
#| warning: false
mag_B_corr = round(compare_data(dt["mag_B"], dt["BT"], sigma = True)[3], 3)
Kmag_corr = round(compare_data(dt["Kmag"], dt["K"], sigma = True)[3], 3)
```

@@

::: panel-tabset
### B mag

@@

### K mag

```{python}
scatter_plot(dt["Kmag"], dt["K"], yerr = dt["E_K"], sigma = True)
```

@@

\[?\]
:::

@@

```{python}
#| echo: false
sfr_data.drop("logSFR_GSW", axis=1, inplace=True)

# Example usage of the function
pairplot_with_correlation(sfr_data)

@@

::: panel-tabset
### Stellar Masses Comparison

@@

dt["diff_M"] =  relative_diff(hec = dt["logM_HEC"].data, lcv = dt["logStellarMass"])
temp = dt[["diff_M"]].to_pandas()
sns.histplot(temp, kde = True)
plt.title("Percentage change Distribution [%]")
plt.xlabel(r"$\log_{10}M_*$ "+f"[{dt['logM_HEC'].unit}]")
plt.show()
plt.close()

@@

```{python}
temp.rename(columns={"diff_M":"Percantage Change [%]"}, inplace = True)
```

### Heatmap

```{python}
mass_data.drop("logM_GSW", axis=1, inplace=True)
# Compute the correlation matrix

pairplot_with_correlation(mass_data)
```
:::
